\documentclass{article}

\usepackage{times}
\usepackage{uist}

\begin{document}

% --- Copyright notice ---
\conferenceinfo{UIST'11}{October 16-19, 2011, Santa Barbara, CA, USA}
\CopyrightYear{2011}
\crdata{978-1-4503-0716-1/11/10}

% Uncomment the following line to hide the copyright notice
\toappear{}
% ------------------------

\bibliographystyle{plain}

\title{Sketch It, Make It: Freehand Drawing for Precision Rapid Fabrication}

\author{
\parbox[t]{9cm}{\centering
	     {\em Author One}\\
	     Institution Name\\
             City, ST, USA\\
	     user@institution.net}
\parbox[t]{9cm}{\centering
	     {\em Author Two}\\
	     Institution Name\\
             City, ST, USA\\
	     user@institution.net}
}

\maketitle

% TODO: change this
\abstract Abstract goes here. Abstract goes here. Abstract goes
here. Abstract goes here. Abstract goes here. Abstract goes
here. Abstract goes here. Abstract goes here. Abstract goes
here. Abstract goes here. Abstract goes here. Abstract goes
here. Abstract goes here. Abstract goes here. Abstract goes here.

\classification{I.3.5 [Computational Geometry and Object Modeling]: Modeling packages}

% TODO: change this
\terms{Design, Human Factors}

\keywords{sketching, rapid fabrication, design tools, constraints}

\tolerance=400 % prevent words from sticking out in the margin

%% \begin{figure}[tb]
%% \vspace{1.9in}
%% \caption{A figure caption.  It is set in 9 point Helvetica type, with a
%% 0.5 cm wider margin on both left and right sides.} 
%% \label{fig-example}
%% \end{figure}

\section{INTRODUCTION}
* Personal fabrication
  - introduce ponoko because it is used later
* Laser cutting
* examples
* why is design for laser cutting hard?
  - tools mismatch
  - allude to Illustrator study
  - common tasks in this little domain:
    + make notches
    + use precise measurements
    + angle and length matching
    + items sometimes must join together
* made tool SIMI to explore solutions to these problems
  - sketch tool. why is this the right paradigm?
  - contributions

\section{RELATED WORK}
Related work on ...

* tools for design (various domains like UI design, graphic design, CAD)
* fabrication
* sketching
* informal interfaces 

\section{USER OBSERVATIONS}

\subsection{Formative Study}

* illustrator
* describe things made in study
* list problems we saw: x y z

\subsection{Ponoko Analysis}

* examined 50 laser-cut objects from ponoko
* common properties from analysis: (examples only)
  - symmetry
  - angle
  - material
  - joint types

\section{SKETCH IT, MAKE IT}

* recall common tasks described in
  - intro
  - formative study
  - ponoko analysis
* lots of screenshots
* map problems from earlier (x y z) to solutions in simi
* overview of interaction

\subsection{Implementation}

* Stylus with offhand button
* Recognizes syntax and gestures
* Syntax: lines, elliptical arcs, splines, circles, ellipses
* Gestures:
  - Latch (3 kinds) 
  - Erase 
  - undo/redo
  - right angle
  - same-length 
  - flow-selection
  - guide points
  - select stencils
* Constraints
  - introduce what they are
  - list types supported: right angle, same length, co-terminate, specific length
  - visual appearance
* Stencils (with or without holes)

\section{EVALUATION}

* screenshots/photos from user study
* other results from user study...

\section{ACKNOWLEDGMENTS}

% TODO: remove
Every paper should cite \cite{sutherland-sketchpad}, if only to placate BibTeX.

% TODO: fill this in later. Leave left blank for blind review.

\bibliography{simi}

\end{document}
