\documentclass[11pt]{now}
\input{packages.tex}

\begin{document}

\isbn{xxxxxxxxxxxxxx}

\DOI{xxxxxxx}

\abstract{Computational support for sketching is an exciting research
  area at the intersection of design research, human-computer
  interaction, and artificial intelligence. Despite the prevalence of
  software tools, most designers begin their work with physical
  sketches. Modern computational tools largely treat design as a
  linear process beginning with a specific problem and ending with a
  specific solution. Sketch based design tools offer another approach
  that may fit design practice better. This article surveys literature
  related to such tools. First, we describe the practical basis of
  sketching---why people sketch, what significance it has in design
  and problem solving, and the cognitive activities it
  supports. Second, we survey computational support for sketching,
  including methods for performing sketch recognition and managing
  ambiguity, techniques for modeling recognizable elements, and
  human-computer interaction techniques for working with
  sketches. Last we propose challenges and opportunities for future
  advances in this field.}

\articletitle{Computational Support for Sketching in Design: A Review}
            

\authorname1{Gabe Johnson}
\affiliation1{Carnegie Mellon University}
\author1address2ndline{Carnegie Mellon University, }
\author1country{USA}
\author1email{johnsogg@cmu.edu}

\authorname2{Mark D. Gross}
\affiliation2{Carnegie Mellon University}
\author2address2ndline{Carnegie Mellon University, }
\author2country{USA}
\author2email{mdgross@cmu.edu}

\authorname3{Jason Hong}
\affiliation3{Carnegie Mellon University}
\author3address2ndline{Carnegie Mellon University, }
\author3country{USA}
\author3email{jasonh@cs.cmu.edu}

\authorname4{Ellen Yi-Luen Do}
\affiliation4{Georgia Institute of Technology}
\author4address2ndline{Georgia Institute of Technology, }
\author4country{USA}
\author4email{ellendo@gatech.edu}

\journal{sample}
\volume{xx}
\issue{xx}
\copyrightowner{xxxxxxxxx}
\pubyear{2008}
            
\maketitle


\cleardoublepage \pagenumbering{roman}

\tableofcontents

\clearpage

\setcounter{page}{0}
\pagenumbering{arabic}

\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{definition}{Definition}[chapter]


\input{review-1-introduction.tex}

\input{review-2-traditional.tex}

\input{review-3-hardware.tex}

\input{review-4-recognition.tex}

\input{review-5-interaction.tex}

\input{review-6-future.tex}

\newpage
\bibliographystyle{nowsort}
\bibliography{sketch-bibliography}

\end{document}
